---
title: "Your Document Title"
author: "Document Author"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---

```{r}
library(tidyverse)
library(knitr)
```

# Exploratory Data Analysis

# Combining the Covid and Mask Policy datasets

```{r}
# covid data
covid.url <- "https://raw.githubusercontent.com/himalayahall/DATA607-FINALPROJECT/master/data/processed/owid/owid-covid-data.csv"
covid <- read_csv(covid.url)

# face covering data
fc.url <- "https://raw.githubusercontent.com/himalayahall/DATA607-FINALPROJECT/master/data/source/face-covering-policies-covid.csv"
fc <- read.csv(fc.url)

# rename columns in fc data to match
# convert date to date
fc <- fc |>
  rename(date = Day, iso_code = Code, location = Entity) |>
  mutate(date = as.Date(date))

# subset the covid data
covid.sub1 <- covid |> select(c(1:4,9,14,49,67))

# combine covid and face covering datasets
co.mask <- covid.sub1 |>
  left_join(fc, by=c("iso_code","location", "date"))

kable(head(co.mask))
```

This data was then written to file and stored in GitHub for easier loading.

# Avg. Mask Rates - Continent

```{r}
# read combined mask data
md <- "https://raw.githubusercontent.com/himalayahall/DATA607-FINALPROJECT/mask-data/data/processed/owid/covid_mask_combined.csv"
co.mask <- read_csv(md)

# avg mask rate by continent by year
co.mask |> na.omit() |>
  mutate(year = strftime(date, "%Y")) |>
  group_by(continent, year) |>
  summarize(avg_mask = mean(facial_coverings)) |>
  ggplot(aes(x = year, y = avg_mask, group=1)) +
  facet_wrap(~continent) +
  geom_col(fill="red") +
  labs(y = "avg", title = "Avg. Mask Rate - Continent")
```

# Avg. Mask Rates - G7

```{r}
# G7 countries
countries <- c("Canada","France","Germany","Italy","Japan","United Kingdom","United States")

# avg mask rate amongst G7 countries
co.mask |> na.omit() |>
  filter(location %in% countries) |>
  mutate(year = strftime(date, "%Y")) |>
  group_by(location, year) |>
  summarize(avg_mask = mean(facial_coverings)) |>
  ggplot(aes(x = year, y = avg_mask, group=1)) +
  facet_wrap(~location) +
  geom_col(fill="red") +
  labs (y = "avg", title = "Avg. Mask Rate - G7 Countries")
```

# Excess Deaths by Avg. Mask Policy

```{r}
# total excess deaths by avg mask policy
co.em <- co.mask |>
  filter(!is.na(excess_mortality_cumulative_per_million)) |>
  select(3,5:10) |>
  rename(excess = excess_mortality_cumulative_per_million,
         face = facial_coverings) 

co.em |> na.omit() |>
  group_by(location) |>
  summarize(total_excess = max(excess), avg_policy = mean(face)) |>
  ggplot(aes(x=avg_policy, y=total_excess)) +
  geom_jitter()
```

# Excess Deaths by Avg. Mask Policy

```{r}
boxplot(excess ~ face, co.em, main="distribution by policy", xlab="policy", ylab="excess deaths")
```

# Statistical Analysis

Does average mask policy predict total excess mortality?

The null hypothesis, $H_0$ - average mask policy *does not* predict total excess mortality. 

The alternative hypothesis, $H_A$ - average mask policy *does* predict total mortality. 

# Relationship Between Mask Policy and Excess Death

```{r message=FALSE}
excess_mask <- co.em |>
  select(1, 6:7) |>
  na.omit() |>
  group_by(location) |>
  summarize(excess = max(excess), policy = mean(face))
  
ggplot(excess_mask, aes(x=policy, y=excess)) +
  geom_jitter() +
  geom_smooth(method = "lm")
```
# Fitting in a Linear Model

```{r}
model.1 <- lm(excess ~ policy, excess_mask)
summary(model.1)
```

# Interpreting the Model

- p-value = 0.0264, significance
- Adjusted R-squared = 0.037, accounts for very little variance

# Residuals

```{r}
ggplot(model.1, aes(x = .resid)) +
  geom_histogram(binwidth=750) +
  xlab("Residuals")
```

# Residuals

```{r}
plot(model.1)
```
